Co-citations in context

### 2019-06-12 cocit_r.tex
edits all over

\subsection{Model Misspecification and the Attributes of Disciplinary Context} \textcolor{blue}{Tandy: I think this is where we can do a statistical test, and show that these numbers are highly unlikely under the random model.} \emph{This is where we want to add the additional data about expected values ~2.5\% for AP as an example.}

Recall our concern regarding model misspecification  in the Uzzi {\em et al.}~model arising from treating all references in a given year as equiprobable, thus allowing substitutions from WoS without consideration of subject matter (as reflected in the publication's discipline) or citation frequency.
Under this model, the probability of replacing a reference by another reference in a given discipline is identical to the proportion of the articles in WoS that are in that particular discipline.
Thus, if the model accurate reflects citation practice, the proportion of references within papers published in a given discipline $D$ would be the same as the proportion of articles in the discipline $D$, and--conversely--the degree to which the proportion deviates from the expected value reflects the degree of model misspecification.
%due to considering disciplinarily irrelevant references as a concern when disciplinary subsets are analyzed using a random graph model that generates expected co-citation frequencies from the WoS superset. 
%A symptom of model misspecification is that random sampling of citations associated with an article do not match with the distribution of citations across the disciplines observed in articles from that discipline. 
We therefore  assembled three exemplar datasets using keyword searches for the terms `applied physics', `immunology', and `metabolism' (Materials and Methods). 
Using the WoS extended classification of 153 subject areas, we then analyzed the disciplinary composition of the references in these three datasets into the five research areas in the Web of Science (life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities) for the years 1985, 1995, and 2005. We observed that 84.2\% of the references in the applied physics dataset were classified as belonging to the WoS physical sciences research area, which is close to the value of 79.5\% for the references in the physical sciences area. Similarly, 93.7\% and 92.1\% of the references in the immunology and metabolism datasets belonged to the life sciences and biomedicine research area, again close to the value of 89.9\% for the references in the life sciences and biomedicine area. \emph{Add in Sitaram's new data}. 
These data indicate strongly that references are likely to belong to the same research area as the publications that cite them. Our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, although that percentage varied among disciplines in the eight specialties they investigated: from approximately 39\% to 89\% for 2006. 
\textcolor{blue}{Would be nice to know the proportion of publications in each of the five major subject areas.}

@@

We predicted that model misspecification, as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies, constraining  reference substitutions to the disciplinary subnetwork would be less than the
model misspecification when substitutions are drawn equiprobably from all of 
the WoS superset. 
The results in Table \ref{tab:kld} indicate that for the set of journals common to both a disciplinary network and the WoS superset,  confirm this prediction: simulations under our model (which constrains substitutions to the
disciplinary subnetwork) consistently have a lower K-L divergence compared to simulations that all replacements uniformly  from the WoS superset. 
Furthermore, the K-L divergence for  the unconstrained model (where the references are equiprobable and allowed to be taken from all of WoS) is generally twice as large as the K-L divergence for the models where the references are constrained to  the disciplinary subnetworks (a ratio of 1.96 to 2.77,  and  greater than 2.0 for eight out of nine cases), demonstrating that simulations that constrain substitutions to the given disciplinary network better model the observed data.

\subsection{Novelty and Conventionality as Determinants of Impact in Disciplinary Contexts}
\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[width=1.5in]{Fig2H01N10.pdf} & \includegraphics[width=1.5in]{Fig2H10N10.pdf} \\
(a) Top 1\% of cited articles & (b) Top 10\% of cited articles \\
\end{tabular}
\caption{Effect of Context on Journal-Pair z-scores and Categorical Hit Rates: Immunology, Applied Physics, and WoS for 1995. Panels (a) and (b) show hit rates for the LNLC, LNHC, HNLC, and HNHC categories for the immunology, applied physics, and WoS datasets when hit articles are defined as the top 1\% and top 10\% of articles, respectively.  Novelty in both panels is defined at the 10th percentile of articles' z-score distributions. The number of data points in the immunology, applied physics, and WoS data sets are 21,917, 18,305, and 476,288, respectively.  The results for the WoS data set mirrored previous results from \cite{uzzi_atypical_2013}, where the highest hit rate was for the HNHC category.  The highest hit rates for the immunology dataset, in contrast, are in the LNHC category. The LNHC category often had the highest hit rate for the immunology dataset for various parameter settings with the HCHN category having the highest hit rate in other cases: metabolism results are similar. The applied physics dataset shows further contrast as the hit rate for it was highest in the HNLC category.  The HNLC category demonstrated the highest hit rate often for applied physics: otherwise the highest hit rate was for HNHC articles.}
\label{fig:Fig2}
\end{figure*}

Figure \ref{fig:Fig2}, Panels (a) and (b), compares hit rates for the four categories among the Immunology, Applied Physics, and WoS datasets for 1995, where the hit rate is defined as the number of hit articles in each category divided by the number of articles in the category. We evaluated the statistical significance of the categorical hit rates using multiple methods, some of which we describe here.  Our first test was based on the null hypotheses that hits were distributed randomly among the four categories with uniform probability in proportion to the number of articles in each category. Using a Chi-Square Goodness of Fit test, rejecting the null hypothesis in favor of the alternate hypothesis supports a non-uniform dispersion of hits: that is, some of the four categories are individually associated with higher than expected, or lower than expected hit rates. The null hypothesis was rejected at a $p<0.001$ in all cases in  Figure \ref{fig:Fig2}, with the exception of the immunology and applied datasets where hit articles are designated as the top 1\% of articles: valid tests were not possible in those instances due to too few expected hits. The null hypothesis was rejected with $p<0.001$ for all valid tests for all parameter settings, all datasets, and all years: hypotheses tests were valid in 73 of 96 instances. We conclude that it is likely that the distribution of hits among categories is not uniform and that instead hit rates vary among the categories in all datasets. 
%(Should we insert a table with all statistical results?)

We computed hit rates for the WoS dataset, which mirrored Uzzi et al.'s results whereby the largest hit rates were for the HNHC category, despite our methodological improvement of sampling citations in proportion to their frequency. We found contrary results in the 1995 immunology dataset where articles in the LNHC category often had the greatest hit rates (Figure \ref{fig:Fig2}).  
Across all years and parameter settings, the highest hit rates in the immunology datasets were sometimes in the LNHC category and sometimes in the HNHC category, and the hit rates for the metabolism datasets reflected this same pattern  (Figure \ref{fig:Fig2}). 
The greatest hit rates for the applied physics data were often in the HNLC category and otherwise in the HNHC category (Figure \ref{fig:Fig2}).
We conclude that Uzzi et al.'s finding of high hit rates in the HNHC category does not hold generally for disciplinary-based datasets and that novel citation patterns are not always indicative of impactful research, as was the case with immunology. 
Furthermore, the categories displaying the greatest hit rate vary with parameter settings and with the year. 
The lack of stable results across parameter settings suggests that parameters must be selected judiciously.

@@

@@

We conclude that high hit rates in the HNHC category do not hold generally for all datasets and that conventionality flavored with novelty is not always indicative of impactful research. 
%The z-scores that change sign across datasets is either contradictory or an acceptable variation due to the difference in reference sets. We contend that it is the former because the inappropriate substitution of citations with those from disciplines that are implausible and because the observed citation frequencies are ignored. 
That fewer z-scores are negative in the WoS dataset relative to the immunology dataset may be directly due to the uniform sampling of references, whereby many resulting journal pairs are never observed in the literature and so the expected frequencies of those observed pairs have lower expected values, thus biasing their z-scores upward. In addition, treating citation as a set versus a multiset means that a random model will sample popular citations downward, thus further increasing their z-scores.\emph{need confirmation from Jim here}.

Contrary to Uzzi et al., we found the category displaying the highest hit rate to be sensitive to the experimental parameter settings, which included the percentage of articles deemed to be hits and the percentile of articles' z-score distributions that delineated between articles of low novelty and high novelty. Faced with this lack of robustness, we are confronted with the necessity of defining defendable parameters, if possible, that determine novelty and conventionality. 
More significantly, we should question whether the simple approach of defining novelty and conventionality as is done this article and other research is sufficient or whether we should expect such variation in the factors associated with impact as we have observed.

>>>>

\subsection{Model Misspecification and the Attributes of Disciplinary Context} 
Recall our concern regarding model misspecification  in the Uzzi {\em et al.}~model arising from treating all references in a given year as equiprobable, thus allowing substitutions from WoS without consideration of subject matter (as reflected in the publication's discipline) or citation frequency.
Under this model, the probability of replacing a reference by another reference in a given discipline is identical to the proportion of the articles in WoS that are in that particular discipline.
Thus, if the model accurate reflects citation practice, the proportion of references within papers published in a given discipline $D$ would be the same as the proportion of articles in the discipline $D$, and--conversely--the degree to which the proportion deviates from the expected value reflects the degree of model misspecification.
%due to considering disciplinarily irrelevant references as a concern when disciplinary subsets are analyzed using a random graph model that generates expected co-citation frequencies from the WoS superset. 
%A symptom of model misspecification is that random sampling of citations associated with an article do not match with the distribution of citations across the disciplines observed in articles from that discipline. 
We therefore  assembled three exemplar datasets using keyword searches for the terms `applied physics', `immunology', and `metabolism' (Materials and Methods). 
Using the WoS extended classification of 153 subject areas, we then analyzed the disciplinary composition of the references in these three datasets into the five research areas in the Web of Science (life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities) for the years 1985, 1995, and 2005. We observed that 84.2\% of the references in the applied physics dataset were classified as belonging to the WoS physical sciences research area, which is close to the value of 79.5\% for the references in the physical sciences area. Similarly, 93.7\% and 92.1\% of the references in the immunology and metabolism datasets belonged to the life sciences and biomedicine research area, again close to the value of 89.9\% for the references in the life sciences and biomedicine area. \emph{Add in Sitaram's new data}. 
These data indicate strongly that references are likely to belong to the same research area as the publications that cite them. Our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, although that percentage varied among disciplines in the eight specialties they investigated: from approximately 39\% to 89\% for 2006.

\textcolor{blue}{I don't think the paragraph above makes the point that the distribution that is observed is inconsistent with the Uzzi et al. model. But to make that point, it would be nice to know the proportion of publications in each of the five major subject areas. With that information,  I think this is where we can do a statistical test, and show that these numbers are highly unlikely under the random model.
In other words, this is where we want to add the additional data about expected values ~2.5\% for AP as an example.}

@@

Since our analyses show substantial model misspecification for the Uzzi {\em et al.} model, we then examined a modification to the model that we conjectured would better fit the observed data.  Specifically, 
we conjectured that model misspecification would be reduced by constraining the substitutions to  disciplinary subnetworks,  and tested this by examining the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and predicted citation distributions,
restricted to  the set of journals in a given disciplinary subnetwork.  
The results (Table \ref{tab:kld}) confirm this prediction: simulations under the constrained model (which constrains substitutions to the
disciplinary subnetwork) consistently have a lower K-L divergence compared to simulations under the unconstrained model of Uzzi {\em et al.},  where all references in WoS for a given year are equiprobable.
Furthermore, the K-L divergence for  the unconstrained model is generally twice as large as the K-L divergence for the models where the references are constrained to  the disciplinary subnetworks (a ratio of 1.96 to 2.77,  and  greater than 2.0 for eight out of nine cases).
These results clearly demonstrate that the modification to the Uzzi {\em et al.} model produced by  constraining reference substitutions to the given disciplinary network better models the observed data.

\subsection{Calculation of Novelty and Conventionality using the constrained model}  
\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[width=1.5in]{Fig2H01N10.pdf} & \includegraphics[width=1.5in]{Fig2H10N10.pdf} \\
(a) Top 1\% of cited articles & (b) Top 10\% of cited articles \\
\end{tabular}
\caption{Effect of Context on Journal-Pair z-scores and Categorical Hit Rates: Immunology, Applied Physics, and WoS for 1995. Panels (a) and (b) show hit rates for the LNLC, LNHC, HNLC, and HNHC categories for the immunology, applied physics, and WoS datasets when hit articles are defined as the top 1\% and top 10\% of articles, respectively.  Novelty in both panels is defined at the 10th percentile of articles' z-score distributions. The number of data points in the immunology, applied physics, and WoS data sets are 21,917, 18,305, and 476,288, respectively.  The results for the WoS data set mirrored previous results from \cite{uzzi_atypical_2013}, where the highest hit rate was for the HNHC category.  The highest hit rates for the immunology dataset, in contrast, are in the LNHC category. The LNHC category often had the highest hit rate for the immunology dataset for various parameter settings with the HCHN category having the highest hit rate in other cases: metabolism results are similar. The applied physics dataset shows further contrast as the hit rate for it was highest in the HNLC category.  The HNLC category demonstrated the highest hit rate often for applied physics: otherwise the highest hit rate was for HNHC articles.}
\label{fig:Fig2}
\end{figure*}

Since the constrained model better fits the observed data, we evaluated the distribution of 
highly cited articles  (i.e., ``hit articles")
 into the four categories (HCHN, LCHN, HCLN, LCLN), for different
thresholds for hit articles.
Figure \ref{fig:Fig2}, Panels (a) and (b), compares hit rates for the four categories among the Immunology, Applied Physics, and WoS datasets for 1995, where the hit rate is defined as the number of hit articles in each category divided by the number of articles in the category. 
The calculation for the hit rates for the WoS dataset (bottom row, Figure \ref{fig:Fig2})
mirrors Uzzi et al.'s results, whereby the largest hit rates were for the HNHC category, despite our methodological improvement of sampling citations in proportion to their frequency. 
However, results observed for the 1995 immunology and metabolism datasets show different
trends, in that  articles in the LNHC category had the greatest hit rates for the top 1\% of cited articles and tied for the top 10\%.
%Across all years and parameter settings, the highest hit rates in the immunology datasets were sometimes in the LNHC category and sometimes in the HNHC category, and the hit rates for the metabolism datasets reflected this same pattern  (Figure \ref{fig:Fig2}). 
On the other hand, the greatest hit rates for the applied physics data were   in the HNLC category for both the top 1\% and top 10\% of all cited articles.  
Thus, the category exhibiting the highest hit rate (among highly cited papers) depends on the
specific disciplinary network, and also to some extent on the threshold for being highly cited.
 %was the case with immunology. 
Furthermore, the categories displaying the greatest hit rate vary to some extent % with parameter settings and 
with the year  \textcolor{blue}{(data not shown)}.
%The lack of stable results across parameter settings suggests that parameters must be selected judiciously.

\textcolor{blue}{I did not review the next two paragraphs}

We evaluated the statistical significance of the categorical hit rates using multiple methods, some of which we describe here.  Our first test was based on the null hypotheses that hits were distributed randomly among the four categories with uniform probability in proportion to the number of articles in each category. Using a Chi-Square Goodness of Fit test, rejecting the null hypothesis in favor of the alternate hypothesis supports a non-uniform dispersion of hits: that is, some of the four categories are individually associated with higher than expected, or lower than expected hit rates. The null hypothesis was rejected at a $p<0.001$ in all cases in  Figure \ref{fig:Fig2}, with the exception of the immunology and applied datasets where hit articles are designated as the top 1\% of articles: valid tests were not possible in those instances due to too few expected hits. The null hypothesis was rejected with $p<0.001$ for all valid tests for all parameter settings, all datasets, and all years: hypotheses tests were valid in 73 of 96 instances. We conclude that it is likely that the distribution of hits among categories is not uniform and that instead hit rates vary among the categories in all datasets. 
%(Should we insert a table with all statistical results?)

@@

\textcolor{blue}{I think we should delete the next paragraph.}
That fewer z-scores are negative in the WoS dataset relative to the immunology dataset may be directly due to the uniform sampling of references, whereby many resulting journal pairs are never observed in the literature and so the expected frequencies of those observed pairs have lower expected values, thus biasing their z-scores upward. In addition, treating citation as a set versus a multiset means that a random model will sample popular citations downward, thus further increasing their z-scores.\emph{need confirmation from Jim here}.

@@

Our study shows strong evidence that a key assumption of the Uzzi {\em et al.} model that  
substitutions of references  are not impacted by discipline deviates from citation practice, and that
constraining the model to a disciplinary subnetwork reduces the model misspecification.
Furthermore,  using the constrained model instead of the unconstrained model to redefine high (or low) novelty and conventionality produces different trends, depending on the subnetwork.
In particular, while Uzzi {\em et al.} found that highly cited papers were most likely to be
in the HNHC category, this observation does not hold when using the improved model.
Instead, unlike Uzzi {\em et al.}, we find that conventionality flavored with novelty is {\em not} generally a feature of impactful research. 
In particular,   high novelty (as defined by Uzzi {\em et al.}) is
%We conclude that Uzzi et al.'s finding of high hit rates in the HNHC category does not hold generally for disciplinary-based datasets and that 
%novel citation patterns are 
not always indicative of impactful research (as reflected in citation counts), as shown by both Immunology and Metabolism having their highest hit rates for low novelty (Figure \ref{fig:Fig2}).
More generally, these results show that the conclusions of Uzzi {\em et al.} of universal trends in highly cited papers
may be the consequence of using a random model that has a poor fit to the observed data.

Of equal concern, and also contrary to Uzzi et al., we found the category displaying the highest hit rate to be sensitive to the percentage of articles deemed to be hit articles (i.e., the definition of highly cited articles), as well as   and the percentile of articles' z-score distributions that delineated between articles of low novelty and high novelty \textcolor{blue}{(data not shown).}
Faced with this lack of robustness, we are confronted with the necessity of defining defendable parameters, if possible, that determine novelty and conventionality. 
More significantly, we should question whether the simple approach of defining novelty and conventionality, as is done in this article and other research, is sufficient or whether we should expect such variation in the factors associated with impact as we have observed.

%The z-scores that change sign across datasets is either contradictory or an acceptable variation due to the difference in reference sets. We contend that it is the former because the inappropriate substitution of citations with those from disciplines that are implausible and because the observed citation frequencies are ignored.
